### 2020-07-08 latex/vnhcs.tex
First draft of the intro to VNHC section plus defs and thms

\begin{defn}
    A set of canonical coordinates \(\{q,p\}\) for a Hamiltonian system 
    \(\mathcal{H}\) are said to be \textit{simply actuated coordinates} if the
    input matrix \(B(q) \in \R^{n \times k}\) is of the form
    \[
        B(q) = \begin{bmatrix}
            \Zmat{(n-k)\times k} \\
            \Id{k}
        \end{bmatrix}
    \]
\end{defn}
\begin{defn}
    A Hamiltonian system is said to be \textit{simply actuated} if there exists
    a canonical transformation from any set of canonical coordinates 
    \(\{q,p\}\) into simply actuated coordinates for \(\mathcal{H}\).
\end{defn}

@@

\section{Virtual Nonholonomic Constraints}
% TODO: Motivation for VNHCs, they are an extension of VHCs, etc.

% TODO: Perform the full development of VNHCs, including its stabilizing
% controller.
\begin{defn}
    A \textit{virtual nonholonomic constraint of order \(k\)} is a relation
    \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is smooth,
    \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all \((q,p) \in h\inv(0)\),
    and there exists a feedback controller \(\tau(q,p)\) stabilizing the set
    \[
        \Gamma = \left\{(q,p) | h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
    \]
    which is called the \textit{constraint manifold}.
\end{defn}
%TODO: Special case: when dM/dqu = 0, show that we have a nice form and that we
%can solve for p_a and the closed-loop dynamics (qu,pu)_dot

>>>>

\begin{defn}
    A set of canonical coordinates \(\{q,p\}\) for a Hamiltonian system 
    \(\mathcal{H}\) are said to be \textit{simply actuated coordinates} if the
    input matrix \(B(q) \in \R^{n \times k}\) is of the form
    \[
        B(q) = \simpleB    
    \]
\end{defn}
\begin{defn}
    A Hamiltonian system is said to be \textit{simply actuated} if there exists
    a canonical transformation from any set of canonical coordinates 
    \(\{q,p\}\) into simply actuated coordinates for \(\mathcal{H}\).
\end{defn}

@@

\section{Virtual Nonholonomic Constraints}
% TODO: Motivation for VNHCs, they are an extension of VHCs, etc.
Let us imagine a child on a swing, who wants to go as high as possible. The
child will push off the ground and start a swing with a very small oscillation,
then extend and retract their feet appropriately to gain energy. 
If a roboticist were designing a machine to do this task, they might design a
control mechanism which makes the legs track a trajectory over time. For
example, they might tell the robot to extend and retract its legs every two
seconds. In ideal situtations, this technique would work perfectly because the
leg motion is synchronized with the swing to gain energy as fast as possible.

Unfortunately for engineers, reality does not provide ideal scenarios very often. 
Most children have an adult pushing the swing to help them go higher; or perhaps
the child is swinging on a windy day. In either case, they adjust their leg
motion accordingly and do not keep track of time when kicking their legs. This
standard control technique of tracking a function of time, known as
\textit{trajectory tracking}, would not work because the disturbance affecting
the swing will desynchronize the motion of the legs with the swing and stop the
energy-gaining effects.

This is a common issue in the control of many biologically-inspired systems. 
One method which can provide more realistic, robust motion is the method of
of virtual holonomic constraints (VHCs). Instead of a robot's actuators tracking
a trajectory over time, VHCs use the actuators to enforce a relation
\(h(q) = 0\) of the configuration \cite{vhcs_for_el_systems}. 
This method has provided incredible results in the development of 
walking robots \cite{vhc_robotic_walking, vhc_stable_walking}, 
vehicle motion \cite{vhc_bicycle, vhc_helicopter}, 
and has even been used to design a snake-like robot that can swim
\cite{vhc_snake}.

Recent papers have attempted to extend these results to use the robot's
actuators to enforce a relation 
\(h(q,\dot{q}) = 0\) of the full state. Since these relations use both
the configuration and generalized velocity, they are called
\textit{nonholonomic} constraints. Furthermore, they are called \textit{virtual}
because the constraint is enforced by the robot's actuators.
We will give our own characterization of these virtual nonholonomic constraints
in the Hamiltonian framework, using the phase \((q,p)\) of a Hamiltonian
system instead of the state \((q,\dot{q})\) of a lagrangian system.

% TODO: Perform the full development of VNHCs, including its stabilizing
% controller.
\begin{defn}
    A \textit{virtual nonholonomic constraint} (VNHC) \textit{of order \(k\)} is a
    relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is
    smooth, \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
    \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
    stabilizing the set
    \[
        \Gamma = \left\{(q,p) \vert h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
    \]
    which is called the \textit{constraint manifold}.
\end{defn}

Defining the error term  \(e = h(q,p)\), stabilizing \(\Gamma\) is equivalent to 
solving for \(\tau(q,p)\) which drives \(e \rightarrow 0\)
and \(\dot{e} \rightarrow 0\).
Let us imagine that we have no further structure on the VNHC. 
Then \(\dot{e} = dh_q \dot{q} + dh_p \dot{p}\), so \(\tau\) appears inside
\(\dot{e}\). To solve for \(\tau\) explicitly, we must have that \(dh_p\) is
invertible, which means that \(h(q,p)\) is strictly monotonic in \(p\).

This is a very restrictive condition, which many VNHCs will not satisfy. For
this reason, it is preferable to have the torque \(\tau\) appear after two 
derivatives of \(e\) so that there is more freedom in the types of constraints
one can use.
The terminology for this is to say that \(e\) is of \textit{relative degree}
\(\{2,2,\ldots,2\}\). We thus define a special type of VNHC which satisfies this
property.

\begin{defn}
    A VNHC of order \(k\) \(h(q,p) = 0\) is \textit{regular} if the output 
    \(e = h(q,p)\) is of relative degree \(\{2,2.\ldots,2\}\) everywhere on the
    constraint manifold \(\Gamma\).
\end{defn}

\citeauthor{hybrid_zero_dynamics_bipedal_nhvcs} showed that relations which use only
the unactuated momentum \(p_u\) result in \(\tau\) appearing only after two
derivatives of \(e\) \cite{hybrid_zero_dynamics_bipedal_nhvcs}. 
In the remainder of this chapter, we will formalize and generalize this idea 
under the following assumption.

% TODO: Write the [p' M p] matrix in our nice nottation
\begin{assm}\label{assm:H-is-simply-actuated}
    The mechanical system under consideration is an underactuated 
    Hamiltonian system of order \((n-k) \geq 0\) in simply
    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\):
    \begin{align*}
        \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q) \\
         &\begin{cases}
            \dot{q} = \Minv p \\
            \dot{p} = -\frac{1}{2} ? - \nabla_q V(q) + \simpleB \tau \\
        \end{cases}
    \end{align*}
\end{assm}

% TODO: write out ??? = [p' dM/(dqu_i)] the matrix of size nxn
\begin{prop}
    A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
    (\ie \(h(q,p) \equiv h(q,p_u)\)) and
    \[
        \rank{\left(dh_q \Minv(q) + dh_{p_u} ??? \right)\simpleB} = k
    \]
    everywhere on the constraint manifold \(\Gamma\).
\end{prop}
\begin{proof}
\end{proof}

%TODO: Special case: when dM/dqu = 0, show that we have a nice form and that we
%can solve for p_a and the closed-loop dynamics (qu,pu)_dot
